---
title: "Advanced SQL"
author: "Mirko Mälicke"
output:
  html_notebook:
    theme: flatly
    toc: yes
  html_document:
    theme: flatly
    toc: yes
  pdf_document:
    toc: yes
---

```{r, echo=FALSE, include=T}
# require the package
if (!("RPostgreSQL" %in% installed.packages())){
  install.packages("RPostgreSQL")
}
if (!("getPass" %in% installed.packages())){
  install.packages("getPass")
}
require(RPostgreSQL)
require(getPass)
require(ggplot2)
library(dplyr)


# establish the connection
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, host='openhydro.de', port=5432, user=getPass('Provide the user'), 
                 password=getPass('Provide the password'), dbname='dwd')
dbListTables(con)
```

Build a temporary table called *name_precipitation*, where you should replace *name* with your actual name. This table shall contain 
all measurements from the data table holding precipitation information. 
```{sql connection=con}
--select * from data limit 10
create temporary table Nele_precipitation as
select id, date, value as precipitation from data where variable_id = 1 
```

Create a view on this table that holds monthly statistics for every station id. Calculate the monthly sum, the highest observed daily 
rainfall sum, the number of extreme rainfalls (daily precipitation > 30).
```{sql connection=con}
create temporary view Nele_prec_summary as
select id, date_trunc('month', date) as month,
  sum(precipitation) as summe,
  max(precipitation) as maxi,
  sum((precipitation > 30.0)::integer) as "extreme rainfall days" 
from Nele_precipitation
group by id, month 
```

Create a view on the last view, that aggregates the same statistics for anual overviews.
```{sql connection=con}
create temporary view Nele_annual_prec as
select id, date_trunc('year', month) as year,
  sum(summe) as summe,
  max(maxi) as maxi,
  sum("extreme rainfall days") as "extreme rainfall days" 
from Nele_prec_summary
group by id, year 
```

Which station observed the highest monthly rainfall sum?
```{sql connection=con}
select n.id, st.name, max(n.summe) as "highest monthly rainfall" from Nele_annual_prec n
join stations st on n.id=st.id
group by n.id, st.name
order by "highest monthly rainfall" desc
limit 1
```

Which station observed the most extreme rainfall occurences within one month? Which within one year?
```{sql connection=con}
select n.id, st.name,  max(n."extreme rainfall days") as "extreme monthly rainfall occurence" from Nele_prec_summary n
join stations st on n.id=st.id
group by n.id, st.name
order by "extreme monthly rainfall occurence" desc
limit 1;
```

```{sql connection=con}
select n.id, st.name,  max(n."extreme rainfall days") as "extreme annual rainfall occurence" from Nele_annual_prec n
join stations st on n.id=st.id
group by n.id, st.name
order by "extreme annual rainfall occurence" desc
limit 1
```

Which was the year with the  most overall extreme rainfall occurences? (ignoring the station)
```{sql connection=con}
--create temporary view Nele_year_most_extreme_rainfalls as
select year, sum("extreme rainfall days") as "extreme rainfall occurences" from Nele_annual_prec 
group by year
order by "extreme rainfall occurences" desc
limit 1
```

Which is the station with the most extreme rainfall events within this year (from the last chunk)? 
Which one with the most extreme rainfall events overall?
```{sql connection=con}
select e.year, e.id, e.name, e."extreme rainfall days" from 
  (select year, sum("extreme rainfall days") as "extreme rainfall occurences" 
  from Nele_annual_prec 
  group by year
  order by "extreme rainfall occurences" desc
  limit 1) y
join
  (select a.id,a.year, s.name, a."extreme rainfall days" from Nele_annual_prec a
  join stations s on s.id=a.id) e
on y.year = e.year
order by e."extreme rainfall days" desc
limit 1
```

```{sql connection=con}
select n.id, s.name, sum(n."extreme rainfall days") as "overall extreme rainfall days" from Nele_annual_prec n
join stations s on n.id=s.id
group by n.id, s.name
order by "overall extreme rainfall days" desc
limit 1
```

Create a view, that includes only the records from _data_ which fulfilled the extreme rainfall constraint.
Join another column to this view, that holds the _type of precipitation_ variable. From this Join, create a histogram of _type of precipitation_. Calculate an overall histogram.
```{sql connection=con}
-- the extreme rainfall constraint: daily precipitation > 30
-- type of precipitation: rskf (name in variables) id = 3

create temporary view Nele_extreme_rainfall as
select id, date, value from data
where variable_id = 1 and value > 30
```

```{sql connection=con, output.var="extreme_rainfall"}
select n.id, n.date, n.value, d.value as type_prec  from Nele_extreme_rainfall n
join data d on d.id=n.id and d.date = n.date
join variables v on v.id = d.variable_id
where v.name = 'type of precipitation'
```

```{sql connection=con, output.var="overall_prec_type"}
select * from data
where variable_id = 3
```


```{r}
extreme_rainfall$extreme <- T
overall <- left_join(overall_prec_type, extreme_rainfall[,c("id", "date", "extreme")])

prec_description <- data.frame(value = c(0,1,4,6,7,8), 
                               descr = c("no precipitation", "only rain (before 1979)", 
                                         "unknown form of recorded precipitation", "only rain", 
                                         "only snow", "rain and snow"))
overall <- left_join(overall, prec_description)
ggplot(extreme_rainfall, aes(type_prec)) + geom_bar() + ggtitle("type of precipitation of extreme rainfall events")
ggplot(overall, aes(descr, fill=extreme)) + geom_bar() + ggtitle("overall type of precipitation")
```


**OPTIONAL:** Can you redo the last step and do the same analysis for 'per station' histograms?
```{sql connection=con, output.var="overall_prec_type1"}
select 
  d1.id, d1.name, 
  d1.date, 
  case when d1.prec > 30 then 'extreme' else 'normal' end as prec, 
  e1.type_prec 
from 
    (select d.id, s.name, d.date, d.value as prec from data d 
    join stations s on d.id=s.id
     where d.variable_id = 1
    ) d1
join 
  (select e.id, e.date, e.value as type_prec from data e where e.variable_id = 3) e1
on d1.id=e1.id and d1.date=e1.date
```

```{r}
overall_stations <- left_join(overall_prec_type1, prec_description, by = c("type_prec"="value"))

part1 <- unique(overall_stations$name)[1:5]
Part1 <- subset(overall_stations, name %in% part1)
ggplot(Part1, aes(descr, fill=prec) ) + geom_bar() + facet_grid(name ~ .)

Part1_extreme <- subset(Part1, prec == "extreme")
ggplot(Part1_extreme, aes(descr)) + geom_bar()  + facet_grid(name ~ .)
```

```{r}
part2 <- unique(overall_stations$name)[6:10]
Part2 <- subset(overall_stations, name %in% part2)
ggplot(Part2, aes(descr, fill=prec) ) + geom_bar() + facet_grid(name ~ .)

Part2_extreme <- subset(Part2, prec == "extreme")
ggplot(Part2_extreme, aes(descr)) + geom_bar()  + facet_grid(name ~ .)
```


Use the insights of the last two chunks and write two sentences about the _type of precipitation_ variable. How useful is it? 
<div class="alert alert-info">
Answer:
The type of precipitation only discriminates between the two types rain and snow. Those two types mostly depend on the temperature, and will not influence the amount of precipitation. Because most extreme rainfalls occur in summer and snow is rare at the most stations all over the year, the result doesn't realy show any informations.
</div>

Load the anual overview on extreme rainfalls into a data.frame and plot a violine plot of the amount of extreme rainfalls per station.
```{sql connection=con, output.var="rain.annual"}
select * from Nele_annual_prec
```
```{r}
#View(rain.annual)
ggplot(rain.annual, aes(factor(id), `extreme rainfall days`)) +
  geom_violin() + xlab("station id") + ylab("amount of extreme rainfall days whithin one year")
```

Redo the last exercise for the monthly extreme rainfall counts. Does this change the picture?
What can you say about the distribution of extreme rainfall events from these plots?
```{sql connection=con, output.var="rain.monthly"}
select * from Nele_prec_summary
```
```{r}
#View(rain.monthly)
ggplot(rain.monthly, aes(factor(id), `extreme rainfall days`)) +
  geom_violin() + xlab("station id") + ylab("amount of extreme rainfall days whithin one month")

head(rain.monthly)
```
<div class="alert alert-info">
Answer:
The plots are different. The amount of extreme rainfall days within one year is in average at least two times as high as the amount within one month. This means that the extreme rainfall events of the year never occured all in one month, even not in a two months time. I would state from this pictures, that the extreme rainfall events are distributet normal over the year.  
</div>

Load the locations for the used stations from the database.
```{sql connection=con, output.var="locations"}
select id, name, st_x(geometry) as lon, st_y(geometry) as lat 
from stations 
where geometry is not null and discipline_id = 1
```
```{r}
View(locations)
ggplot(locations, aes(x=lon, y=lat)) + geom_point(shape=18)
```


Plot these locations and represent any measure for annual variance in extreme rainfall events by the point size of the used locations.<br>
**OPTIONAL:** Additionally mark the station showing the most and least recorded extreme rainfalls by another marker into the same plot. (eg. use a green and a red cross or star.)
```{r}
rain.locations <- left_join(rain.annual, locations, by="id")
rain.var <- rain.locations %>% 
  group_by(id, name, lon, lat) %>% 
  summarise(var= var(`extreme rainfall days`), sum.extreme.rainfalls = sum(`extreme rainfall days`)) %>% 
# add a column that specifies the extremes:
  mutate(extreme = ifelse(sum.extreme.rainfalls == max(rain.var$sum.extreme.rainfalls),
                          "max", ifelse(sum.extreme.rainfalls == min(rain.var$sum.extreme.rainfalls), "min", "medium")))
ggplot(rain.var, aes(x=lon, y=lat)) + 
  geom_point(aes(size=var, color=extreme)) 
```

How can you interpret this result?
<div class="alert alert-info">
Answer:
The stations with big variance in extreme rainfall events have had some years with many events and some with few. They therefore have a heterogenity over the years. The stations with little variance are affected by the extreme rainfall events mostly in the same amount, and as the colors show, have only had few rainfall events at all. Probably lots of extreme rainfall events are rare, but sometimes they occur. The rare occurence of years with lots of extreme rainfall events makes up the high variance. One might therefore expect a correlation between an increasing variance of yearly extreme rainfall events and the increasing overall amount of extreme rainfall events.
</div>

### cleanup
```{r, include=FALSE}
dbDisconnect(con)
```




