---
title: "SQL Exercises"
author: "Mirko Mälicke"
output: html_notebook
---

This notebook collects some basic SQL exercises for the sql_introduction.Rmd lecture. You can solve them and create a pull request for your soultion in the master Github branch.

```{r, include=FALSE}
# require the package
if (!("RPostgreSQL" %in% installed.packages())){
  install.packages("RPostgreSQL")
}
require(RPostgreSQL)
if (!("getPass" %in% installed.packages())){
  install.packages("getPass")
}
require(RPostgreSQL)
require(getPass)

# establish the connection
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, host='openhydro.de', port=5432, user=getPass('Provide the user'), 
                 password=getPass('Provide the password'), dbname='dwd')
```


### SELECT

#### get an overview:
```{r}
tabellen <- dbListTables(con)
```


Select all stations located lower than 400 meter.
```{sql connection=con}
select name from stations where elevation < 400
```

How many stations start with the letter 'W'?
```{sql connection=con}
select count(name) from stations where name like 'W%' or name like 'w%'
```

Are there more stations starting with 'M' or 'W'?
```{sql connection=con}
select substr(name, 1, 1) as first_letter, count(name) as count from stations 
where (name like 'M%' or name like 'W%' )
group by first_letter 
order by count desc limit 1
```

Are the 'M' or the 'W' stations higher in mean?
```{sql connection=con}
--select * from stations limit 7
select substr(name, 1, 1) as first_letter, avg(elevation) from stations 
where name like 'M%' or name like 'W%' or name like 'm%' or name like 'w%'
group by first_letter
order by avg(elevation) desc
```

### SELECT & Grouping

Find the 5 starting letters in station names with the most occurences.
```{sql connection=con}
select substr(name, 1, 1) as first5, count(name) as name_occ from stations
group by first5
order by name_occ desc
limit 5
```

Find the 10 starting letters in station names with the lowest mean elevation.
```{sql connection=con}
select substr(name, 1, 1) as first10, avg(elevation) from stations
group by first10
order by avg(elevation) asc
limit 10
```

You also learned about the vegetation_cover table. The *description_id* attribute was highly repetitive. But how does the distribution look like between these labels? Count the occurences per description_id.

```{sql connection=con}
select description_id, count(description_id) from vegetation_cover
group by description_id
```

Group the vegetation_cover table by the name in vegetation_cover_description and count the instances per label.

```{sql connection=con}
select   count(v.id),   d.name from vegetation_cover as v
  join vegetation_cover_description as d
  on  d.id = v.description_id
  group by d.name
```


### cleanup
```{r, include=FALSE}
dbDisconnect(con)
```

